% $Id: appendix-b.tex 269 2003-04-22 14:45:04Z mackers $

\setlength{\parindent}{0em}

\chapter{Project Diary}

\subsubsection{Week Starting November 4th}

Background reading of W3C guidelines and reports and other web
tutorials. Wrote draft of ``introduction'' and ``background'' chapters.

\subsubsection{Week Starting November 11th}

Started to create sample set to get broad idea of technologies
involved.

Investigated related work.

\subsubsection{Week Starting December 20th}

Created first ``requirements and specifications'' draft.

\subsubsection{Week Starting January 6th}

Designed project architecture and class structure

\subsubsection{Week Starting January 13th}

Implemented skeletal classes.

\subsubsection{January 21st}

Read the Swing tutorial, played with sample applications and created
first SWJFrame window.

\subsubsection{January 23rd}

Investigated internationalisation features. Created sw.properties file.
Replaced menu names with internationalised versions.

\subsubsection{January 24th}

Played with java Xerces. Added XML/DOM parsing function to XHTMLImporter.
Added temporary ``Debug'' menu to program.

\subsubsection{February 7th}

Preliminary DOM parsing. Code to map table rows and cells to internal
data structures. Summary and caption parsing.

\subsubsection{February 10th}

Worked on table display classes. Implemented \code{getComponent} method
in cells and header cells.

\subsubsection{February 16th}

Implemented \code{NullCell} to add row and column spanning support. Algorithms
for filling cells correctly around spanned cells and short rows. Aligned
with W3C specifications.

\subsubsection{February 17th}

Added cell selection support using a mouse listener on the table. Updated
cell class to change background colour for selected cells.

\subsubsection{February 18th}

Placed tables in scroll pane to allow user to scroll big tables. Added status
bar with placeholder text. Added window title bar text.

\subsubsection{February 20th}

Implemented id/abbr/axis and basic headers importing and hooked this in to
the status bar display code. Table switching now available from table menu.

\subsubsection{March 12th}

Implemented HTML importing using JTidy, a CSV importer and a XHTML exporter
using JTidy's pretty print function.

\subsubsection{March 13th}

Improved header import algorithm to support scopes and theads. Also imported
the header finder algorithm for unclaimed cells. 

Implemented `AI' header finder algorithm using heuristics for detecting
headers marked up as bold, etc.

Spanned cells are now displayed prettily in the user interface and the ``no-mans-land''
is painted white.

Improved exporter to export existing cell attributes and correct tag names
for guessed headers.

\subsubsection{March 14th}

Work on menus; added separators and context enabling based on selected cells, 
as per requirements.

Misc. testing and bug fixing and usability improvements with dialog boxes for
import errors such as nested/missing table testing. Added filters to import
dialog box. Added view source functionality.

Added pre-parsers via ``generator'' regular expression.

\subsubsection{March 15th}

JAR file testing revealed some bugs. Testing on Windows revealed more bugs.

\subsubsection{March 18th}

Dialog day - implemented user interface and functionality of all dialogs, which 
was simply a matter of calling the appropriate methods on the existing document,
table and cell classes.

\subsubsection{March 30th}

Added descriptions to dialogs.

\subsubsection{April 3rd}

Added mnemonics to menus and menu items.

\subsubsection{April 4th}

Added remaining Swing accessibility requirements, such as giving components
accessible names and descriptions where necessary and linking labels to their
related components.

\subsubsection{April 6th}

Added any remaining javadoc comments.
